---
title: "Lab 9: Baby Names"
format:
  html:
    self-contained: true
editor: visual
author: "Luke Thilmony"
---

## Setup

```{r}
#| message: false
library(tidyverse)
```

```{r}
#| message: false
names_a <- read_csv(here::here('supporting_artifacts', 'Learning Targets', 'StateNames_A.csv'))
```

## 3 Summarizing and Visualizing

### 1.

```{r}
names_allison <- names_a |> 
  filter(
    Name == 'Allison'
  ) |> 
  group_by(
    State,
    Gender
  ) |> 
  summarize(sum = sum(Count)) |> 
  pivot_wider(
    names_from = Gender,
    values_from = sum
  ) |> 
  mutate(
    M = replace_na(M, 0)
  ) |> 
  rename(
    "Number of Female Babies" = `F`,
    "Number of Male Babies" = M
  )

names_allison
```

### 2.

```{r}
names_allison <- names_allison |> 
  select(
    State,
    `Number of Female Babies`
  )
```

### 3.

```{r}
allison_years <- names_a |> 
  filter(
    Name == 'Allison'
  ) |> 
  group_by(
    Year
  ) |> 
  summarize(
    sum = sum(Count)
  )

allison_years |> ggplot(
  mapping = aes(x = Year, y = sum)
) + 
  geom_line() +
  labs(
    y = element_blank(),
    title = 'Frequency of Babies Named "Allison"'
  )
```

## 4 Modeling the Number of Allisons

### 4.

```{r}
allison_model <- allison_years |> 
  lm(
    sum ~ Year,
    data = _
  )
```

### 5.

```{r}
allison_model |> 
  ggplot(
    mapping = aes(x = Year, y = sum)
  ) +
  geom_point() +
  stat_smooth(method = 'lm') +
  labs(
    y = 'Number of Babies Named Allison'
  )
```

### 6.

predicted count<sub>allison</sub> = 209815.1 - 101.6(Year)

### 7.

```{r}
allison_model |> 
  broom::augment() |> 
  ggplot(
    mapping = aes(y = .resid, x = .fitted)
  ) +
  geom_point() +
  labs(
    x = 'Predicted',
    title = 'Residual Plot',
    y = 'Residuals'
  ) +
  geom_hline(
    yintercept = 0,
    linetype = 'dashed'
  )
```

There are no clear trends in the residual plot but there is a group of predicted counts from around 5900 to 6600 that were all underestimated.

### 8.

Your name is declining in coolness/popularity.

## 5 Spelling by State

### 1.

```{r}
names_allan <- names_a |> 
  filter(
    Name == 'Allan' | Name == 'Alan' | Name == 'Allen'
  )
```

### 2.

```{r}
names_allan <- names_allan |> 
  filter(
    State == 'CA' | State == 'PA',
    Year == '2000'
  ) |> 
  group_by(
    State,
    Name
  ) |> 
  summarize(
    sum = sum(Count)
  ) |> 
  pivot_wider(
    names_from = Name,
    values_from = sum
  )

names_allan
```

### 3.

```{r}
names_allan_pct <- names_allan |> 
  mutate(
    total = sum(Alan, Allan, Allen),
    Alan = (Alan / total) * 100,
    Allan = (Allan / total) * 100,
    Allen = (Allen / total) * 100
  ) |> 
  select(
    -total
  )

names_allan_pct
```

In CA, 65.54% were named Alan, 14.7% were named Allan, and 19.75% were named Allen. In PA, 42.86% were named Alan, 10.08% were named Allan, and 47.06% were named Allen.
