---
title: "Lab 4: Avocado Prices"
format: html
editor: visual
author: "Luke Thilmony"
---

## Setup

```{r Setup and Ingest}
#| message: false
library(tidyverse)

avocados <- read_csv(
  here::here(
    "supporting_artifacts",
    "avocado.csv"
  )
)
```

## Description

### 1.

The dataset contains data from the national retail volume and price of Hass avocados from 2013 to 2018. It contains the date of the observation, the average price of one avocado, the type (organic or conventional), the year, the region of the observation, the total number of avocados sold, and three variables Part Lookup Codes that describe the size of avocado sold.

## Cleaning

### 2.

```{r cleaning}
avocados_clean <- avocados |> 
  group_by(
    region
  ) |>
  rename(
    small = `4046`,
    large = `4225`,
    extra_large = `4770`
  )

regions <- data.frame(
  region = c(
    "GreatLakes",
    "Northeast",
    "Midsouth",
    "NorthernNewEngland",
    "Plains",
    "SouthCentral",
    "Southeast",
    "West",
    "WestTexNewMexico"
  )
)

states <- data.frame(
  region = c(
    "California",
    "SouthCarolina",
    "NewYork"
  )
)

totalUS <- data.frame(
  region = c(
    "TotalUS"
  )
)

avocados_region <- avocados_clean |> 
  semi_join(
    y = regions,
    by = "region"
  ) 

avocados_states <- avocados_clean |>
  semi_join(
    y = states,
    by = "region"
  )

avocados_metro <- avocados_clean |> 
  anti_join(
    y = regions,
    by = "region"
  ) |> 
  anti_join(
    y = states,
    by = "region"
  ) |> 
  anti_join(
    y = totalUS,
    by = "region"
  )
```

## Exercises

### 3.

```{r region}
avocados_region <- avocados_region |> 
  group_by(
    region
  ) |> 
  select(
    small,
    type,
    year,
    region
  ) |> 
  filter(
    year == 2017,
    type == "organic"
  ) |> 
  mutate(
    sum = sum(small)
  ) |> 
  distinct(
    region,
    sum
  ) |> 
  ungroup() |> 
  filter(
    sum == max(sum)
  )

avocados_region
```

The region with the most small avocados sold in 2017 was the West region.

### 4.

```{r month}
avocados_clean |> 
  separate(
    col = Date,
    into = c("year", "month", "day"),
    sep = "-"
  ) |> 
  anti_join(
    totalUS,
    by = 'region'
  ) |> 
  group_by(
    month
  ) |>
  mutate(
    sum = sum(`Total Volume`)
  ) |> 
  distinct(
    month,
    sum
  ) |> 
  ungroup() |> 
  filter(
    sum == max(sum)
  )
```

The month with the highest avocado sales is February.

### 5.

```{r metro}
total_metro <- avocados_metro |> 
  group_by(
    region
  ) |> 
  select(
    `Total Volume`,
    region
  ) |> 
  mutate(
    sum = sum(`Total Volume`)
  ) |> 
  distinct(
    region,
    sum
  ) |> 
  arrange(
    sum
  ) |> 
  tail(
    n = 5
  )

total_metro
```

Los Angeles sold the most avocados in terms of total volume.

```{r metro_plot}
top5 <- data.frame(
  region = c(
    'Denver',
    'Houston',
    'PhoenixTuscon',
    'DallasFtWorth',
    'LosAngeles'
  )
)

avocados_metro |> 
  semi_join(
    top5,
    by = 'region'
  ) |> 
  ggplot(
    mapping = aes(x = `Total Volume`, y = region),
  ) +
  geom_boxplot() +
  labs(
    y = 'Region',
    title = 'Top Five Regions with the Highest Average Volume Avocados Sold'
  )
```

## Reshaping

### 6.

```{r CA_cities}
ca_cities <- data.frame(
  region = c(
    'LosAngeles',
    'SanDiego',
    'Sacramento',
    'SanFrancisco'
  )
)

ca_av <- avocados_metro |> 
  semi_join(
    ca_cities,
    by = 'region'
  )
```

```{r CA_cities_analysis}
ca_av |> 
  group_by(
    region,
    type
  ) |> 
  mutate(
    mean_avg = mean(AveragePrice)
  ) |> 
  distinct(
    region,
    .keep_all = TRUE
  ) |>
  select(
    region,
    type,
    mean_avg
  ) |> 
  pivot_wider(
    names_from = type,
    values_from = mean_avg
  ) |> 
  mutate(
    diff = abs(conventional - organic)
  )

ca_av |> 
  group_by(
    region,
    type
  ) |> 
  select(
    region,
    type,
    AveragePrice
  ) |> 
  ggplot(
    mapping = aes(
      x = AveragePrice,
      y = region,
      col = type
    )
  ) +
  geom_boxplot() +
  labs(
    title = "Average Price of Avocados in California",
    x = "Average Price per Avocado (USD)",
    y = "Region of California",
    col = 'Type'
  )
  
```

The region with the highest difference between the average cost of types of avocados is San Francisco. The mean price for an organic avocado was \$2.21 dollars while the mean price for a conventional avocado was \$1.39, a difference of 0.814 dollars.

### 7.

```{r plotting}
ca_av_plotting <- ca_av |> 
  select(
    `Total Volume`,
    small,
    large,
    extra_large,
    type,
    region
  ) |> 
  group_by(
    region,
    type
  ) |> 
  mutate(
    total = sum(`Total Volume`),
    small_prop = sum(small) / total,
    large_prop = sum(large) / total,
    xlarge_prop = sum(extra_large) / total
  ) |> 
  distinct(
    region,
    type,
    .keep_all = TRUE
  ) |> 
  select(
    type:xlarge_prop
  ) |> 
  pivot_longer(
    cols = small_prop:xlarge_prop,
    names_to = "size"
  )

ggplot(
  data = ca_av_plotting,
  mapping = aes(
    fill = factor(size, levels = c('small_prop', 'large_prop', 'xlarge_prop')),
    x = region,
    y = value
  )
) +
  geom_col(position = 'fill') +
  facet_wrap(~ type) +
  labs(
    y = 'Proportion of Mean Avocados Sold',
    x = 'Region of CA',
    fill = 'Avocado Size',
  ) +
  scale_fill_discrete(labels = c('Small', 'Large', 'Extra Large')) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

# https://stackoverflow.com/questions/47025828/change-the-order-of-stacked-fill-columns-in-ggplot2
# for the reordering of stacking

# https://datavizpyr.com/how-to-dodge-overlapping-text-on-x-axis-labels-in-ggplot2/
# for the scale_x_discrete specifics
```
