---
title: "Lab 7: Functions and Fish"
format:
  html:
    self-contained: true
editor: visual
author: "Luke Thilmony"
---

## The Data

```{r}
#| message: false

library(tidyverse)
```

```{r ingest}
#| message: false

fish <- read_csv(here::here("supporting_artifacts", "BlackfootFish.csv"))
```

## Part One: Summaries and Plots

### Task 1 -- Summarizing Missing Data

```{r}
obs_na <- fish |> 
  summarize(
    across(
      .cols = everything(),
      .fns = ~ sum(is.na(.))
    )
  )
  
obs_na
```

There are 1796 observations with missing data. The variable with all of the missing data points is weight.

### Task 3 -- Visualizing Missing Data

```{r}
#| fig-align: 'center'
fish_plots <- fish |> 
  group_by(
    year, 
    trip, 
    section
  ) |> 
  summarize(
    across(
      .cols = everything(),
      .fns = ~ sum(is.na(.))
    )
  ) |> 
  select(
    year,
    trip,
    section,
    weight
  ) |> 
  mutate(
    trip = as.factor(trip)
  )

fish_plots |> ggplot(
    mapping = aes(
      x = year, y = weight, fill = section
    )
  ) + 
  geom_col() +
  facet_grid(~ fct_recode(trip, `Trip 1` = '1', `Trip 2` = '2')) +
  labs(
    title = 'Frequency of Mission Data Over Time by Section',
    subtitle = 'Faceted by Trip',
    x = 'Year',
    y = element_blank(),
    fill = 'Section'
  )
```

## Part Two: Adjusting the Data

### Task 1 & 2-- Writing a Function, Adding Stops

```{r}
rescale01 <- function(vec) {
  stopifnot(
    is.numeric(vec),
    length(vec) != 1
  )
  
  vec_min <- min(vec, na.rm = T)
  vec_max <- max(vec, na.rm = T)
  
  
  res <- (vec - vec_min) / (vec_max - vec_min)
  return(res)
}
```

### Task 3 -- Simple Test

```{r}
x <- c(1:25, NA)
rescale01(x)
max(rescale01(x), na.rm = T)
min(rescale01(x), na.rm = T)
```

### Task 4 -- Performing a More Difficult Test

```{r}
#| message: false
#| layout-ncol: 2
norescale <- ggplot(
  data = fish,
  mapping = aes(x = length)
)

norescale +
  geom_boxplot() +
  labs(
    title = 'No Length Rescaling'
  )
norescale +
  geom_histogram() +
  labs(
    title = 'No Length Rescaling'
  )

rescale <- ggplot(
  data = fish,
  mapping = aes(x = rescale01(length))
) 
rescale +
  geom_boxplot()+
  labs(
    title = 'Rescaled Length'
  )
rescale +
  geom_histogram() +
  labs(
    title = 'Rescaled Length'
  )

```

### Task 5 -- Incorporating Variables

```{r}
rescale_columns <- function(df, var) {
  stopifnot(is.data.frame(df))
  
  res <- df |> 
    mutate(
      across(
        .col = {{var}},
        .fns = rescale01
      )
    )
  
  return(res)
}
```

### Task 6 -- Another Function Test

```{r}
res <- rescale_columns(fish, c(length, weight))
res
```
